Lambda Calculus through Tropical Mathematics

### 2023-01-09 main.tex
Nuove modifiche

\title{Tropical Geometry and the Lambda Calculus}

@@

\section{Two Quantitative Approaches to the $\lambda$-calculus}
\input{section2}

\section{Tropical Linear Algebra and Linear Programs}
\input{section3}

\section{Tropical Polynomials and Non-Linear Programs  }
\input{section4}

@@

\section{Related Work}
\input{section8}

>>>>

\title{Lambda Calculus through Tropical Mathematics}

@@

\section{Two Quantitative Approaches to the $\lambda$-calculus}

\input{section2}

\section{Tropical (non-)Linear Algebra and computation}

Di cosa stiamo parlando. Ovvero:

- matrices over a continuous semi-ring

- The tropical semi-ring $\Lawv$ coincides with the Lawvere quantale.

- tropical linear algebra: the category $\LREL$

\subsection{As a model of linear programs}

\input{section3}

\subsection{As a model of non-linear programs}

\input{section4}

\subsection{As a model of differential programs}

- differential categories

- the tropical differential operator: formal differentiation of tropical Laurent series. Differentiation in $\LREL_{!}$.

- $\LREL_{!}$ is a differential CCC satisfying Taylor. 
Discuss the tropical Taylor expansion $\trop\C T(M)$.

@@

\section{Related Work (vale la pena? O meglio nelle conclusioni?)}
\input{section8}

### 2023-01-09 section2.tex
Nuove modifiche

Recall the two approaches with more details on lambda-calculus and on existing challenges.

@@

@@

- resource lambdas are polynomials, so not linear. Why we need tropical geometry here.

>>>>

%Recall the two approaches with more details on lambda-calculus and on existing challenges.

\subsection{Metric approach: bounded $\lambda$-terms as Lipschitz functions}

@@

\subsection{Resource approach: differential $\lambda$-terms as polynomials}

@@

- resource lambdas are polynomials, so not linear.

\subsection{Tropical mathematics: a possible synthesis}

- generalities about tropical maths (tropicalisation $\trop P$  of polynomials and of Laurent series, and their roots -- all that without $\LREL$)

### 2023-01-09 section4.tex
Nuove modifiche

@@

- tropical polynomials and tropical roots -> bounded lambda-terms
(the operation $\C M_{\leq n}(-)$ defines a graded linear exponential comonad) via the adjunction
$$
\LREL(Z\otimes !_{\leq n}X,Y) \simeq \LREL(Z, !_{\leq n}X\multimap Y) 
$$

- tropical Laurent series and tropical roots -> unbounded lambda-terms (the operation $\C M(-)$ defines a linear exponential comonad) via the adjunction
$$
\LREL(!(Z\& X),Y) \simeq \LREL(!Z, !X\multimap Y) 
$$

- tropicalization of analytic functions and their roots: 
$\trop P$

- the differential operator: formal differentiation of tropical Laurent series. Differentiation in $\LREL_{!}$.

- $\LREL_{!}$ is a differential CCC satisfying Taylor. 
Discuss the tropical Taylor expansion $\trop\C T(M)$.

@@

>>>>

\subsubsection{Bounded}

Quanto vogliamo essere precisi in questa sottosottosezione ?

@@

@@

- tropical polynomials and tropical roots $->$ bounded lambda-terms
(the operation $\C M_{\leq n}(-)$ defines a graded linear exponential comonad) via the adjunction
$$
\LREL(Z\otimes !_{\leq n}X,Y) \simeq \LREL(Z, !_{\leq n}X\multimap Y) 
$$

\subsubsection{Unbounded}

- tropical Laurent series and tropical roots $->$ unbounded lambda-terms (the operation $\C M(-)$ defines a linear exponential comonad) via the adjunction
$$
\LREL(!(Z\& X),Y) \simeq \LREL(!Z, !X\multimap Y) 
$$

That is:

- $\LREL_!$ is CCC (Giulio)
